---
title: "Untitled"
author: "Adrian Figueroa Rodriguez"
date: "25 de octubre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Corse Description
## Conditionals and Control Flow
### Equality
```{r}
3 == (2+1)
"intermediate" != "r"
TRUE != FALSE
"Rchitect" != "rchitect"
# All these expressiones are true!

# Other Instructions
TRUE == FALSE
-6*14 != 17-101
"useR" == "user"
TRUE == 1

```

#### Greater and less than

```{r}
(1+2) > 4
"dogs"< "Cats"
TRUE <= FALSE

# comparison of numerics
-6*5+2 >= -10+1
# comparison caracters
"raining" <= "raining dogs"
TRUE > FALSE
```

### Compare vectors

```{r}
linkedin <- c(16,9,13,5,2,17,14)
facebook <- c(17,7,5,16,8,13,14)

linkedin > 15
linkedin <= 5
```

### compare matrices
```{r}
linkedin <- c(16,9,13,5,2,17,14)
facebook <- c(17,7,5,16,8,13,14)

views <- matrix(c(linkedin,facebook), nrow = 2, byrow = TRUE)
views

# when does views = 13

views == 13
views <= 14
```

## Logical Operators
### & and |

```{r}
TRUE & TRUE
FALSE | TRUE
5 <= 5 & 7 <6 
3 < 4 | 7 < 6

# EXAMPLE
linkedin <- c(16,9,13,5,2,17,14)
last <- tail(linkedin,1)
5 > last | last > 10
15 < last & last <= 20
```
### & and | (2)

```{r}
linkedin
facebook
views
# linked exceeds 10 but facebook beleow 10
linkedin > 10 & facebook <10

linkedin >= 12 | facebook >= 12

views > 11 & views <= 14
```

### Blend it all together

```{r}
# Read data
#li_df <- read.table(stdin(), header = TRUE)

# Writing data
# Write CSV in R
# write.csv(li_df, file = "data/li_df.csv")

#li_df <- read.csv("data/li_df.csv")

#second <- li_df$day2
#second

#extremes <- second >25 | second < 5
#extremes

#sum(extremes)

#sum(li_df$days2 > 25 | li_df$day2 < 5)

```
## Conditional statements
### The if statement

```{r}
medium = "LinkedIn"
num_views = 15
if(medium == "LinkedIn"){
  print("Showing LinkedIn infromation")
}

if(num_views > 15){
  print("You're popular!")
}
```

### Add an else
```{r}
medium <- "LinkedIn"
num_views <- 14

if(medium == "LinkedIn"){
  print("Showing LinkedIn information")
} else{
  print("Unkown medium")
}

if(num_views > 15){
  print("You're popular!")
} else{
  print("Try to be more visible!")
}
```
### Customize further: else if

```{r}
if(medium == "LinkedIn"){
  print("Showing LinkedIn information")
} else if(medium == "Facebook"){
  print("Showing Facebook info")
} else {
  print("Unkown medium")
}

if(num_views > 15){
  print("You're popular!")
} else if(num_views <= 15 & num_views > 10){
  print("Your numer of views is average ")
} else{
  print("Try to be more visible!")
}
```
### Else if 2.0
```{r}
number <- 5

if(number <10){
  if(number < 5){
    result <- "Extra small"
  } else{
    result <- "Small"
  }
} else if(number < 100){
  result <- "medium"
} else{
  result <- "large"
}

print(result)
```

### Take Control!
```{r}
li <- 15
fb <- 9

if (li >= 15 & fb >= 15){
  sms <- 2 * (li + fb)
} else if(li < 10 & fb < 10){
  sms <- 0.5 * (li + fb)
} else{
  sms <- li + fb
}

sms

```

## Loops
### Write a while loop
```{r}
speed <- 64

while (speed > 30){
  print("Slow down!")
  speed <- speed -7
}

speed

```

### Throw in more conditionals
```{r}
speed <- 64

while(speed > 29){
  print(paste("Your speed is", speed))
  
  if(speed > 48){
    print("Slow down big time!")
    speed <- speed - 11
    
  } else if(speed > 30){
    print("Slow down!")
    speed <- speed - 6
    
  } else if(speed == 30){
    print("Ok, you're at a ok speed")
    speed <- speed - 2
  }
}
```

### Stop the while loop: break
```{r}

speed <- 88

while (speed > 30){
  print(paste("Your speed is", speed))
  if(speed > 80){
    break
  }
  if(speed >40){
    print("Slow down big time!")
    speed <- speed - 11
  } else{
    print("Slow down")
    speed <- speed - 6
  }
}
```

### Build a while loop from scratch
```{r}
i <- 1

while (i <= 10){
  print (3 * i )
  if((3*i) %% 8 == 0){
    break
  }
  i <- i+1
}
```
### Loop over a vector
```{r}
primes <- c(2,3,5,7,11,13)

for(p in primes){
  print(p)
}

for(i in 1:length(primes)){
  print(primes[i])
}


linkedin <- c(16, 9, 13, 5, 2, 17, 14)

for(li in linkedin){
  print(li)
}

# my try

for(lnk in linkedin){
  print(lnk)
}

for(i in 1:length(linkedin)){
  print(linkedin[i])
}

```

### Loop over a list

```{r}
primes_list <- list(2,3,5,7,11,13)

for(p in primes_list){
  #My try
  if(p/2 > 2){
    print(p)  
  }
}

nyc <- list(pop = 8405837, boroughs = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"), capital = FALSE)

for(info in nyc){
  print(info)
}

for(i in 1:length(nyc)){
  print(nyc[i])
}
```




### Loop over a matrix

```{r}
#for(var1 in seq1){
#  for(var2 in seq2){
    #expression()
#  }
#}

d1 <- c("O",  "  ",   "X")
d2 <- c("  ", "O",  "O" )
d3 <- c("X",  "  ",  "X")

ttt_data <- c(d1,d2,d3)
ttt_data

ttt <- matrix(ttt_data, byrow = TRUE, nrow = 3)
ttt

for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
    print(paste("On row", i, "and column", j, "the board contains", ttt[i,j]))
  }
}
```

### Mix it up with control flow

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

for(li in linkedin){
  if(li > 10){
    print("You're popular!")
  } else{
    print("Be more visible!")
  }
  print(li)
}
```

### Next, you break it

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

for(li in linkedin){
  if(li > 10){
    print("You're popular!")
  } else {
    print("Be more visible!")
  }

if(li > 16){
  print("This is ridiculous, I'm outta here!")
  break
}

if(li < 5){
  print("This is too embarrasing!")
  next()
}

print(li)
}

```

### Build a for loop from scratch

```{r}
rquote <- "r's internals are irrefutably intriguing"
rquote

chars <- strsplit(rquote, split = "")[[1]]
chars

rcount <- 0

for(char in chars){
  if(char == "r"){
    rcount <- rcount + 1
  }
  if(char == "u"){
    break
  }
}

rcount
```

## Functions
### Function documentation

```{r}
help(sample)
args(sample)
args(remove)
help(mean)
args(mean)

```

### Use a function

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

avg_li <- mean(x = linkedin)
avg_fb <- mean(facebook)

avg_li
avg_fb
```

### Use a function
```{r}
# mean(x, ...)
# mean(x, trim = 0, na.rm = FALSE, ...)

linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

avg_sum <- mean(linkedin + facebook)

avg_sum_trimmed <- mean(linkedin + facebook, trim = 0.2)

avg_sum

avg_sum_trimmed
```

### Use a function(3)
```{r}
# mean(x, trim = 0, na.rm = FALSE, ...)

linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

mean(linkedin)

mean(linkedin, na.rm = TRUE)
```

### Functions inside functions
```{r}
speed <- 31
print(paste("Your speed is", speed))

linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

mean(abs(linkedin - facebook), na.rm = TRUE)
```

### Write your own function
```{r}
my_fun <- function(arg1, arg2){
  x <- (arg1 * arg2) + 1 
  print(x)
}

my_fun(11,12)

pow_two <- function(x){
  x ^ 2
}

pow_two(12)

sum_abs <- function(x,y){
  abs(x) + abs(y)
}

sum_abs(-2,3)
```

### Write your own function (2)

```{r}
throw_die <- function() {
  number <- sample(1:6, size = 1)
  number
}

throw_die()

hello <- function() {
  print("Hi there!")
  TRUE
}

hello()

```


### Write your own function (3)

```{r}
my_func <- function(arg1, arg2){
  x <- arg1 + arg2
  x
}

my_func(1,4)


pow_two <- function(x, print_info = TRUE){
  y <- x ^ 2
  if(print_info){
    print(paste(x, "to the power two equals", y))
  }
  return(y)
}  

pow_two(5)
pow_two(5, FALSE)
```

### R you functional?

```{r}

linkedin <- c(16, 9, 13,  5, 2, 17, 14)
facebook <- c(17, 7, 5, 16,  8, 13, 14)

interpret <- function(num_views) {
  if (num_views > 15){
    print("You're popular!")
    return(num_views)
    
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

interpret(linkedin[1])
interpret(facebook[2])
```

### R you functional? (2)

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

interpret_all <- function(views, return_sum = TRUE) {
  count <- 0

  for (v in views) {
    count <- count + interpret(v)
  }

  if (return_sum) {
    return(count)
  } else {
    return(NULL)
  }
}

interpret_all(linkedin)
interpret_all(facebook)
```
### Load an R Package
```{r}
library("ggplot2")

qplot(mtcars$wt, mtcars$hp)
```
## The apply family
### Use lapply with a built-in R function

```{r}
#lapply()
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

split_math <- strsplit(pioneers, split = ":")
split_math

split_low <- lapply(split_math, tolower)
split_low 

str(split_low)
```
### Use lapply with your own function
```{r}
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

split <- strsplit(pioneers, split = ":")
split

split_low <- lapply(split, tolower)
split_low

select_first <- function(x) {
  x[1]
}

names <- lapply(split_low, select_first)
names

select_second <- function(x) {
  x[2]
}

years <- lapply(split_low, select_second)
years
```
### lapply and anonymous functions
```{r}
triple <- function(x) { 3 * x }

function(x) { 3 * x }

lapply(list(1,2,3), function(x) { 3 * x })

pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

names <- lapply(split_low, function(x) { x[1] })
names

years <- lapply(split_low, function(x) { x[2] })
years
```

### Use lapply with additional arguments
```{r}
multiply <- function(x, factor) {
  x * factor
}
lapply(list(1,2,3), multiply, factor = 3)

pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

select_el <- function(x, index) {
  x[index]
}

names <- lapply(split_low, select_el, index = 1)
names

years <- lapply(split_low, select_el, index = 2)
years
```

























 